Music Perception 2: Analysis

### 2022-01-31 experiment-2/analysis/analysis.Rmd
finalise analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Music Perception 2: Analysis"
output: html_notebook
---

@@

- 2 blocks (articulation / dynamics): participants were asked to do the same task for two techniques (expressions) separately. The order of the blocks was counterbalanced across participants.
- 36 trials for each block (each stimulus was presented only once) and the stimuli were randomly presented across participants.

@@

3. Correct answers would be 1) participants pressed "YES" when they listened to teaching performances or 2) participants pressed "NO" when they listened to performing performances.

@@

```{r packages, include = FALSE}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# midi
if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
if (!require("qqplotr")) {install.packages("qqplotr"); require("qqplotr")}
# stats
if (!require("ppcor")) {install.packages("ppcor"); require("ppcor")}
if (!require("performance")) {install.packages("performance"); require("performance")}
```

```{r read, include = FALSE}
# read files and combine them
data_ls <- list.files("data", pattern = "csv")
combined <- lapply(data_ls, function(f){
  fread(paste("data/", f, sep = ""), sep = ",")
})
all_data <- do.call(rbind.data.frame, combined)

@@

```{r midi, include = FALSE}
# create a list of data file names
lf <- list.files("../experiment/mid/", pattern = "mid")

# combine data
all_data <- data.table()
for (file in lf){
   current <- readMidi(paste("../experiment/mid/", file, sep = ""))
   current$MidFile <- gsub(".mid", "", file)
   all_data <- rbind(all_data, current)
}

data <- all_data[event == "Note On" | event == "Note Off"]
data$NoteNr <- rep(1:144, nrow(data)/144)
data$channel <- NULL
data$parameterMetaSystem <- NULL
data$track <- NULL
data$type <- NULL
data$time2 <- (data$time/960)*1000 # deltatime to time (ms)

# labelling
colnames(data)[c(3,4)] <- c("Pitch", "Velocity") # time event also change
data$Condition <- "NA"
data$Skill <- "NA"
data[grepl("_t_", MidFile)]$Condition <- "teaching"
data[grepl("_p_", MidFile)]$Condition <- "performing"
data[grepl("a_", MidFile)]$Skill <- "articulation"
data[grepl("d_", MidFile)]$Skill <- "dynamics"

@@

# assign Subcomponents
dt_ioi$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_ioi[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_ioi[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_ioi[Skill == "dynamics" & Interval == ls_forte[[phrase]][note]]$Subcomponent <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_ioi[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
 }
}

@@

ggscatter(ioi, x = "Mean", y = "Teaching", color = "Skill", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "IOIs (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```
## Articulation (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(ioi[Skill == "articulation"]$Teaching)
shapiro.test(ioi[Skill == "articulation"]$Mean)

cor.test(ioi[Skill == "articulation"]$Teaching, ioi[Skill == "articulation"]$Mean)
```

## Dynamics (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(ioi[Skill == "dynamics"]$Teaching)
shapiro.test(ioi[Skill == "dynamics"]$Mean)

cor.test(ioi[Skill == "dynamics"]$Teaching, ioi[Skill == "dynamics"]$Mean)

cor.test(ioi[Skill == "dynamics"]$Teaching, ioi[Skill == "dynamics"]$Mean, method = "spearman", exact = FALSE)
```

@@

## Articulation (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(ioi_tra[Skill == "articulation"]$Teaching)
shapiro.test(ioi_tra[Skill == "articulation"]$Mean)

@@

## Dynamics (*)
```{r, echo = FALSE}
# normality check
shapiro.test(ioi_tra[Skill == "dynamics"]$Teaching)
shapiro.test(ioi_tra[Skill == "dynamics"]$Mean)

@@

ggscatter(cv, x = "CV", y = "Teaching", color = "Skill", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```
## Articulation (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(cv[Skill == "articulation"]$Teaching)
shapiro.test(cv[Skill == "articulation"]$CV)

@@

## Dynamics (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(cv[Skill == "dynamics"]$Teaching)
shapiro.test(cv[Skill == "dynamics"]$CV)

@@

ggscatter(kot_all, x = "Mean", y = "Teaching", color = "Subcomponent", add = "reg.line", facet.by = "Skill",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = FALSE,
          xlab = "KOT (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```
## Legato (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(kot_all[Subcomponent == "Legato"]$Teaching)
shapiro.test(kot_all[Subcomponent == "Legato"]$Mean)

cor.test(kot_all[Subcomponent == "Legato"]$Teaching, kot_all[Subcomponent == "Legato"]$Mean)
```

## Staccato (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(kot_all[Subcomponent == "Staccato"]$Teaching)
shapiro.test(kot_all[Subcomponent == "Staccato"]$Mean)

cor.test(kot_all[Subcomponent == "Staccato"]$Teaching, kot_all[Subcomponent == "Staccato"]$Mean)
```

## Forte (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(kot_all[Subcomponent == "Forte"]$Teaching)
shapiro.test(kot_all[Subcomponent == "Forte"]$Mean)

cor.test(kot_all[Subcomponent == "Forte"]$Teaching, kot_all[Subcomponent == "Forte"]$Mean)

cor.test(kot_all[Subcomponent == "Forte"]$Teaching, kot_all[Subcomponent == "Forte"]$Mean, method = "spearman", exact = FALSE)
```

## Piano (*)
```{r, echo = FALSE}
# normality check
shapiro.test(kot_all[Subcomponent == "Piano"]$Teaching)
shapiro.test(kot_all[Subcomponent == "Piano"]$Mean)

cor.test(kot_all[Subcomponent == "Piano"]$Teaching, kot_all[Subcomponent == "Piano"]$Mean)

cor.test(kot_all[Subcomponent == "Piano"]$Teaching, kot_all[Subcomponent == "Piano"]$Mean, method = "spearman", exact = FALSE)
```

@@

## Forte (**)
```{r, echo = FALSE}
# normality check
shapiro.test(vel_all[Subcomponent == "Forte"]$Teaching)
shapiro.test(vel_all[Subcomponent == "Forte"]$Mean)

cor.test(vel_all[Subcomponent == "Forte"]$Teaching, vel_all[Subcomponent == "Forte"]$Mean)

cor.test(vel_all[Subcomponent == "Forte"]$Teaching, vel_all[Subcomponent == "Forte"]$Mean, method = "spearman", exact = FALSE)
```

## Piano (*)
```{r, echo = FALSE}
# normality check
shapiro.test(vel_all[Subcomponent == "Piano"]$Teaching)
shapiro.test(vel_all[Subcomponent == "Piano"]$Mean)

cor.test(vel_all[Subcomponent == "Piano"]$Teaching, vel_all[Subcomponent == "Piano"]$Mean)

cor.test(vel_all[Subcomponent == "Piano"]$Teaching, vel_all[Subcomponent == "Piano"]$Mean, method = "spearman", exact = FALSE)
```
## Legato (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(vel_all[Subcomponent == "Legato"]$Teaching)
shapiro.test(vel_all[Subcomponent == "Legato"]$Mean)

cor.test(vel_all[Subcomponent == "Legato"]$Teaching, vel_all[Subcomponent == "Legato"]$Mean)
```

## Staccato (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(vel_all[Subcomponent == "Staccato"]$Teaching)
shapiro.test(vel_all[Subcomponent == "Staccato"]$Mean)

cor.test(vel_all[Subcomponent == "Staccato"]$Teaching, vel_all[Subcomponent == "Staccato"]$Mean)
```

@@

```{r vel-diff, include = FALSE}
data_onset$Diff <- diff(c(0, data_onset$Velocity))
dt_vel_diff <- data_onset[NoteNr != 1]
# assign Interval
dt_vel_diff$Interval <- rep(1:71, nrow(dt_vel_diff)/71)

# for intervals
dt_vel_diff$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
for (note in 1:length(ls_legato[[phrase]])){
  dt_vel_diff[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
}
}
# Staccato
for (phrase in 1:length(ls_staccato)){
for (note in 1:length(ls_staccato[[phrase]])){
  dt_vel_diff[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
}
}
# Forte
for (phrase in 1:length(ls_forte)){
for (note in 1:length(ls_forte[[phrase]])){
  dt_vel_diff$Subcomponent[dt_vel_diff$Skill == "dynamics" & dt_vel_diff$Interval == ls_forte[[phrase]][note]] <- "Forte"
}
}
# Piano
for (phrase in 1:length(ls_piano)){
for (note in 1:length(ls_piano[[phrase]])){
  dt_vel_diff[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
}
}
# Assign Skill Change
for (i in change_1){
dt_vel_diff[Skill == "articulation" & Interval == i]$Subcomponent <- "LtoS"
dt_vel_diff[Skill == "dynamics" & Interval == i]$Subcomponent <- "FtoP"
}
for (i in change_2){
dt_vel_diff[Skill == "articulation" & Interval == i]$Subcomponent <- "StoL"
dt_vel_diff[Skill == "dynamics" & Interval == i]$Subcomponent <- "PtoF"
}
```

@@

## Forte to Piano (***)
```{r, echo = FALSE}
# normality check
shapiro.test(vel_diff_all[Subcomponent == "FtoP"]$Teaching)
shapiro.test(vel_diff_all[Subcomponent == "FtoP"]$Mean)

cor.test(vel_diff_all[Subcomponent == "FtoP"]$Teaching, vel_diff_all[Subcomponent == "FtoP"]$Mean)

cor.test(vel_diff_all[Subcomponent == "FtoP"]$Teaching, vel_diff_all[Subcomponent == "FtoP"]$Mean, method = "spearman", exact = FALSE)
```

## Piano to Forte (***)
```{r, echo = FALSE}
# normality check
shapiro.test(vel_diff_all[Subcomponent == "PtoF"]$Teaching)
shapiro.test(vel_diff_all[Subcomponent == "PtoF"]$Mean)

cor.test(vel_diff_all[Subcomponent == "PtoF"]$Teaching, vel_diff_all[Subcomponent == "PtoF"]$Mean)

cor.test(vel_diff_all[Subcomponent == "PtoF"]$Teaching, vel_diff_all[Subcomponent == "PtoF"]$Mean, method = "spearman", exact = FALSE)
```
## Legato to Staccato (n.s.)
```{r, echo = FALSE}
# normality check
shapiro.test(vel_diff_all[Subcomponent == "LtoS"]$Teaching)
shapiro.test(vel_diff_all[Subcomponent == "LtoS"]$Mean)

cor.test(vel_diff_all[Subcomponent == "LtoS"]$Teaching, vel_diff_all[Subcomponent == "LtoS"]$Mean)
```

## Staccato to Legato (*)
```{r, echo = FALSE}
# normality check
shapiro.test(vel_diff_all[Subcomponent == "StoL"]$Teaching)
shapiro.test(vel_diff_all[Subcomponent == "StoL"]$Mean)

cor.test(vel_diff_all[Subcomponent == "StoL"]$Teaching, vel_diff_all[Subcomponent == "StoL"]$Mean)
```

@@

# KOT
## Legato (n.s.)
```{r kot-partial-leg, echo = TRUE}
pcor.test(partial[Subcomponent == "Legato"]$KOT, partial[Subcomponent == "Legato"]$Teaching, partial[Subcomponent == "Legato", c("IOI", "KV", "KVDiff")])
```

>>>>

---
title: "Music Perception 2: Analysis"
output:
  html_document:
    df_print: paged
---

@@

- 2 blocks (articulation / dynamics): each block only contained performances with either of articulation or dynamics. The order of the blocks was counterbalanced across participants.
- 36 trials for each block (each stimulus was presented only once) and the stimuli were randomly presented within the block.

@@

@@

```{r setup, include = FALSE}
options(knitr.duplicate.label = "allow")

# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# midi
if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
if (!require("corrplot")) {install.packages("corrplot"); require("corrplot")}
if (!require("qqplotr")) {install.packages("qqplotr"); require("qqplotr")}
# stats
if (!require("ppcor")) {install.packages("ppcor"); require("ppcor")}
if (!require("performance")) {install.packages("performance"); require("performance")}
```

```{r read, include = FALSE}
# read files and combine them
data_lf <- list.files("../data", pattern = "csv")
combined <- lapply(data_lf, function(f){
  fread(paste("../data/", f, sep = ""), sep = ",")
})
all_data <- do.call(rbind.data.frame, combined)

@@

```{r midi, include = FALSE}
# create a list of data file names
midi_lf <- list.files("../experiment/mid/", pattern = "mid")

# combine data
all_midi <- data.table()
for (file in midi_lf){
   current <- readMidi(paste("../experiment/mid/", file, sep = ""))
   current$MidFile <- gsub(".mid", "", file)
   all_midi <- rbind(all_midi, current)
}

midi <- all_midi[event == "Note On" | event == "Note Off"]
midi$NoteNr <- rep(1:144, nrow(midi)/144)
midi$channel <- NULL
midi$parameterMetaSystem <- NULL
midi$track <- NULL
midi$type <- NULL
midi$time2 <- (midi$time/960)*1000 # deltatime to time (ms)

# labelling
colnames(midi)[c(3,4)] <- c("Pitch", "Velocity") # time event also change
midi$Condition <- "NA"
midi$Skill <- "NA"
midi[grepl("_t_", MidFile)]$Condition <- "teaching"
midi[grepl("_p_", MidFile)]$Condition <- "performing"
midi[grepl("a_", MidFile)]$Skill <- "articulation"
midi[grepl("d_", MidFile)]$Skill <- "dynamics"

@@

# assign Subcomponents
dt_ioi$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_ioi[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_ioi[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
 }
}
# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_ioi[Skill == "dynamics" & Interval == ls_forte[[phrase]][note]]$Subcomponent <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_ioi[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
 }
}

@@

ggscatter(ioi, x = "Mean", y = "Teaching", color = "Skill", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "IOIs (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

## Articulation (n.s.)
```{r, echo = FALSE}
# normality check
ioi_art_norm_teaching <- shapiro.test(ioi[Skill == "articulation"]$Teaching)
ioi_art_norm_judge <- shapiro.test(ioi[Skill == "articulation"]$Mean)

qqnorm(ioi[Skill == "articulation"]$Teaching)
qqline(ioi[Skill == "articulation"]$Teaching)
ioi_art_norm_teaching

qqnorm(ioi[Skill == "articulation"]$Mean)
qqline(ioi[Skill == "articulation"]$Mean)
ioi_art_norm_judge

cor_ioi_art <- cor.test(ioi[Skill == "articulation"]$Teaching, ioi[Skill == "articulation"]$Mean)
cor_ioi_art
```

## Dynamics (n.s.)
```{r, echo = FALSE}
# normality check
ioi_dyn_norm_teaching <- shapiro.test(ioi[Skill == "dynamics"]$Teaching)
ioi_dyn_norm_judge <- shapiro.test(ioi[Skill == "dynamics"]$Mean)

qqnorm(ioi[Skill == "dynamics"]$Teaching)
qqline(ioi[Skill == "dynamics"]$Teaching)
ioi_dyn_norm_teaching

qqnorm(ioi[Skill == "dynamics"]$Mean)
qqline(ioi[Skill == "dynamics"]$Mean)
ioi_dyn_norm_judge

cor_ioi_dyn <- cor.test(ioi[Skill == "dynamics"]$Teaching, ioi[Skill == "dynamics"]$Mean)
cor_ioi_dyn

cor_ioi_dyn_spearman <- cor.test(ioi[Skill == "dynamics"]$Teaching, ioi[Skill == "dynamics"]$Mean, method = "spearman", exact = FALSE)
cor_ioi_dyn_spearman
```

@@

## Articulation (n.s.)
```{r, echo = FALSE}
# normality check
ioi_art_tra_norm_teaching <- shapiro.test(ioi_tra[Skill == "articulation"]$Teaching)
ioi_art_tra_norm_judge <- shapiro.test(ioi_tra[Skill == "articulation"]$Mean)

qqnorm(ioi_tra[Skill == "articulation"]$Teaching)
qqline(ioi_tra[Skill == "articulation"]$Teaching)
ioi_art_tra_norm_teaching

qqnorm(ioi_tra[Skill == "articulation"]$Mean)
qqline(ioi_tra[Skill == "articulation"]$Mean)
ioi_art_tra_norm_judge

@@

## Dynamics (*)
```{r, echo = FALSE}
# normality check
ioi_dyn_tra_norm_teaching <- shapiro.test(ioi_tra[Skill == "dynamics"]$Teaching)
ioi_dyn_tra_norm_judge <- shapiro.test(ioi_tra[Skill == "dynamics"]$Mean)

qqnorm(ioi_tra[Skill == "dynamics"]$Teaching)
qqline(ioi_tra[Skill == "dynamics"]$Teaching)
ioi_dyn_tra_norm_teaching

qqnorm(ioi_tra[Skill == "dynamics"]$Mean)
qqline(ioi_tra[Skill == "dynamics"]$Mean)
ioi_dyn_tra_norm_judge

@@

ggscatter(cv, x = "CV", y = "Teaching", color = "Skill", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

## Articulation (n.s.)
```{r, echo = FALSE}
# normality check
cv_art_norm_teaching <- shapiro.test(cv[Skill == "articulation"]$Teaching)
cv_art_norm_judge <- shapiro.test(cv[Skill == "articulation"]$CV)

qqnorm(cv[Skill == "articulation"]$Teaching)
qqline(cv[Skill == "articulation"]$Teaching)
cv_art_norm_teaching

qqnorm(cv[Skill == "articulation"]$CV)
qqline(cv[Skill == "articulation"]$CV)
cv_art_norm_judge

@@

## Dynamics (n.s.)
```{r, echo = FALSE}
# normality check
cv_dyn_norm_teaching <- shapiro.test(cv[Skill == "dynamics"]$Teaching)
cv_dyn_norm_judge <- shapiro.test(cv[Skill == "dynamics"]$CV)

qqnorm(cv[Skill == "dynamics"]$Teaching)
qqline(cv[Skill == "dynamics"]$Teaching)
cv_dyn_norm_teaching

qqnorm(cv[Skill == "dynamics"]$CV)
qqline(cv[Skill == "dynamics"]$CV)
cv_dyn_norm_judge

@@

ggscatter(kot_all, x = "Mean", y = "Teaching", color = "Subcomponent", add = "reg.line", facet.by = "Skill",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = FALSE,
          xlab = "KOT (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

## Legato (n.s.)
```{r, echo = FALSE}
# normality check
kot_leg_norm_teaching <- shapiro.test(kot_all[Subcomponent == "Legato"]$Teaching)
kot_leg_norm_judge <- shapiro.test(kot_all[Subcomponent == "Legato"]$Mean)

qqnorm(kot_all[Subcomponent == "Legato"]$Teaching)
qqline(kot_all[Subcomponent == "Legato"]$Teaching)
kot_leg_norm_teaching

qqnorm(kot_all[Subcomponent == "Legato"]$Mean)
qqline(kot_all[Subcomponent == "Legato"]$Mean)
kot_leg_norm_judge

cor_kot_leg <- cor.test(kot_all[Subcomponent == "Legato"]$Teaching, kot_all[Subcomponent == "Legato"]$Mean)
cor_kot_leg
```

## Staccato (n.s.)
```{r, echo = FALSE}
# normality check
kot_sta_norm_teaching <- shapiro.test(kot_all[Subcomponent == "Staccato"]$Teaching)
kot_sta_norm_judge <- shapiro.test(kot_all[Subcomponent == "Staccato"]$Mean)

qqnorm(kot_all[Subcomponent == "Staccato"]$Teaching)
qqline(kot_all[Subcomponent == "Staccato"]$Teaching)
kot_sta_norm_teaching

qqnorm(kot_all[Subcomponent == "Staccato"]$Mean)
qqline(kot_all[Subcomponent == "Staccato"]$Mean)
kot_sta_norm_judge

cor_kot_sta <- cor.test(kot_all[Subcomponent == "Staccato"]$Teaching, kot_all[Subcomponent == "Staccato"]$Mean)
cor_kot_sta
```

## Forte (n.s.)
```{r, echo = FALSE}
# normality check
kot_for_norm_teaching <- shapiro.test(kot_all[Subcomponent == "Forte"]$Teaching)
kot_for_norm_judge <- shapiro.test(kot_all[Subcomponent == "Forte"]$Mean)

qqnorm(kot_all[Subcomponent == "Forte"]$Teaching)
qqline(kot_all[Subcomponent == "Forte"]$Teaching)
kot_for_norm_teaching

qqnorm(kot_all[Subcomponent == "Forte"]$Mean)
qqline(kot_all[Subcomponent == "Forte"]$Mean)
kot_for_norm_judge

cor_kot_for <- cor.test(kot_all[Subcomponent == "Forte"]$Teaching, kot_all[Subcomponent == "Forte"]$Mean)
cor_kot_for

cor_kot_for_spearman <- cor.test(kot_all[Subcomponent == "Forte"]$Teaching, kot_all[Subcomponent == "Forte"]$Mean, method = "spearman", exact = FALSE)
cor_kot_for_spearman
```

## Piano (*)
```{r, echo = FALSE}
# normality check
kot_pia_norm_teaching <- shapiro.test(kot_all[Subcomponent == "Piano"]$Teaching)
kot_pia_norm_judge <- shapiro.test(kot_all[Subcomponent == "Piano"]$Mean)

qqnorm(kot_all[Subcomponent == "Piano"]$Teaching)
qqline(kot_all[Subcomponent == "Piano"]$Teaching)
kot_pia_norm_teaching

qqnorm(kot_all[Subcomponent == "Piano"]$Mean)
qqline(kot_all[Subcomponent == "Piano"]$Mean)
kot_pia_norm_judge

cor_kot_pia <- cor.test(kot_all[Subcomponent == "Piano"]$Teaching, kot_all[Subcomponent == "Piano"]$Mean)
cor_kot_pia

cor_kot_pia_spearman <- cor.test(kot_all[Subcomponent == "Piano"]$Teaching, kot_all[Subcomponent == "Piano"]$Mean, method = "spearman", exact = FALSE)
cor_kot_pia_spearman
```

@@

## Forte (**)
```{r, echo = FALSE}
# normality check
vel_for_norm_teaching <- shapiro.test(vel_all[Subcomponent == "Forte"]$Teaching)
vel_for_norm_judge <- shapiro.test(vel_all[Subcomponent == "Forte"]$Mean)

qqnorm(vel_all[Subcomponent == "Forte"]$Teaching)
qqline(vel_all[Subcomponent == "Forte"]$Teaching)
vel_for_norm_teaching

qqnorm(vel_all[Subcomponent == "Forte"]$Mean)
qqline(vel_all[Subcomponent == "Forte"]$Mean)
vel_for_norm_judge

cor_vel_for <- cor.test(vel_all[Subcomponent == "Forte"]$Teaching, vel_all[Subcomponent == "Forte"]$Mean)
cor_vel_for

cor_vel_for_spearman <- cor.test(vel_all[Subcomponent == "Forte"]$Teaching, vel_all[Subcomponent == "Forte"]$Mean, method = "spearman", exact = FALSE)
cor_vel_for_spearman
```

## Piano (*)
```{r, echo = FALSE}
# normality check
vel_pia_norm_teaching <- shapiro.test(vel_all[Subcomponent == "Piano"]$Teaching)
vel_pia_norm_judge <- shapiro.test(vel_all[Subcomponent == "Piano"]$Mean)

qqnorm(vel_all[Subcomponent == "Piano"]$Teaching)
qqline(vel_all[Subcomponent == "Piano"]$Teaching)
vel_pia_norm_teaching

qqnorm(vel_all[Subcomponent == "Piano"]$Mean)
qqline(vel_all[Subcomponent == "Piano"]$Mean)
vel_pia_norm_judge

cor_vel_pia <- cor.test(vel_all[Subcomponent == "Piano"]$Teaching, vel_all[Subcomponent == "Piano"]$Mean)
cor_vel_pia

cor_vel_pia_spearman <- cor.test(vel_all[Subcomponent == "Piano"]$Teaching, vel_all[Subcomponent == "Piano"]$Mean, method = "spearman", exact = FALSE)
cor_vel_pia_spearman
```
## Legato (n.s.)
```{r, echo = FALSE}
# normality check
vel_leg_norm_teaching <- shapiro.test(vel_all[Subcomponent == "Legato"]$Teaching)
vel_leg_norm_judge <- shapiro.test(vel_all[Subcomponent == "Legato"]$Mean)

qqnorm(vel_all[Subcomponent == "Legato"]$Teaching)
qqline(vel_all[Subcomponent == "Legato"]$Teaching)
vel_leg_norm_teaching

qqnorm(vel_all[Subcomponent == "Legato"]$Mean)
qqline(vel_all[Subcomponent == "Legato"]$Mean)
vel_leg_norm_judge

cor_vel_leg <- cor.test(vel_all[Subcomponent == "Legato"]$Teaching, vel_all[Subcomponent == "Legato"]$Mean)
cor_vel_leg
```

## Staccato (n.s.)
```{r, echo = FALSE}
# normality check
vel_sta_norm_teaching <- shapiro.test(vel_all[Subcomponent == "Staccato"]$Teaching)
vel_sta_norm_judge <- shapiro.test(vel_all[Subcomponent == "Staccato"]$Mean)

qqnorm(vel_all[Subcomponent == "Staccato"]$Teaching)
qqline(vel_all[Subcomponent == "Staccato"]$Teaching)
vel_sta_norm_teaching

qqnorm(vel_all[Subcomponent == "Staccato"]$Mean)
qqline(vel_all[Subcomponent == "Staccato"]$Mean)
vel_sta_norm_judge

cor_vel_sta <- cor.test(vel_all[Subcomponent == "Staccato"]$Teaching, vel_all[Subcomponent == "Staccato"]$Mean)
cor_vel_sta
```

@@

```{r vel-diff, include = FALSE}
data_onset$Diff <- diff(c(0, data_onset$Velocity))
dt_vel_diff <- data_onset[NoteNr != 1]
dt_vel_diff$Interval <- rep(1:71, nrow(dt_vel_diff)/71)

# assign Subcomponents
dt_vel_diff$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
  for (note in 1:length(ls_legato[[phrase]])){
    dt_vel_diff[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
  }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
  for (note in 1:length(ls_staccato[[phrase]])){
    dt_vel_diff[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
  }
}
# Forte
for (phrase in 1:length(ls_forte)){
  for (note in 1:length(ls_forte[[phrase]])){
    dt_vel_diff$Subcomponent[dt_vel_diff$Skill == "dynamics" & dt_vel_diff$Interval == ls_forte[[phrase]][note]] <- "Forte"
  }
}
# Piano
for (phrase in 1:length(ls_piano)){
  for (note in 1:length(ls_piano[[phrase]])){
    dt_vel_diff[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
  }
}

# assign Subcomponent Change
for (i in change_1){
  dt_vel_diff[Skill == "articulation" & Interval == i]$Subcomponent <- "LtoS"
  dt_vel_diff[Skill == "dynamics" & Interval == i]$Subcomponent <- "FtoP"
}
for (i in change_2){
  dt_vel_diff[Skill == "articulation" & Interval == i]$Subcomponent <- "StoL"
  dt_vel_diff[Skill == "dynamics" & Interval == i]$Subcomponent <- "PtoF"
}
```

@@

## Forte to Piano (***)
```{r, echo = FALSE}
# normality check
vel_diff_ftop_norm_teaching <- shapiro.test(vel_diff_all[Subcomponent == "FtoP"]$Teaching)
vel_diff_ftop_norm_judge <- shapiro.test(vel_diff_all[Subcomponent == "FtoP"]$Mean)

qqnorm(vel_diff_all[Subcomponent == "FtoP"]$Teaching)
qqline(vel_diff_all[Subcomponent == "FtoP"]$Teaching)
vel_diff_ftop_norm_teaching

qqnorm(vel_diff_all[Subcomponent == "FtoP"]$Mean)
qqline(vel_diff_all[Subcomponent == "FtoP"]$Mean)
vel_diff_ftop_norm_judge

cor_vel_diff_ftop <- cor.test(vel_diff_all[Subcomponent == "FtoP"]$Teaching, vel_diff_all[Subcomponent == "FtoP"]$Mean)
cor_vel_diff_ftop

cor_vel_diff_ftop_spearman <- cor.test(vel_diff_all[Subcomponent == "FtoP"]$Teaching, vel_diff_all[Subcomponent == "FtoP"]$Mean, method = "spearman", exact = FALSE)
cor_vel_diff_ftop_spearman
```

## Piano to Forte (***)
```{r, echo = FALSE}
# normality check
vel_diff_ptof_norm_teaching <- shapiro.test(vel_diff_all[Subcomponent == "PtoF"]$Teaching)
vel_diff_ptof_norm_judge <- shapiro.test(vel_diff_all[Subcomponent == "PtoF"]$Mean)

qqnorm(vel_diff_all[Subcomponent == "PtoF"]$Teaching)
qqline(vel_diff_all[Subcomponent == "PtoF"]$Teaching)
vel_diff_ptof_norm_teaching

qqnorm(vel_diff_all[Subcomponent == "PtoF"]$Mean)
qqline(vel_diff_all[Subcomponent == "PtoF"]$Mean)
vel_diff_ptof_norm_judge

cor_vel_diff_ptof <- cor.test(vel_diff_all[Subcomponent == "PtoF"]$Teaching, vel_diff_all[Subcomponent == "PtoF"]$Mean)
cor_vel_diff_ptof

cor_vel_diff_ptof_spearman <- cor.test(vel_diff_all[Subcomponent == "PtoF"]$Teaching, vel_diff_all[Subcomponent == "PtoF"]$Mean, method = "spearman", exact = FALSE)
cor_vel_diff_ptof_spearman
```
## Legato to Staccato (n.s.)
```{r, echo = FALSE}
# normality check
vel_diff_ltos_norm_teaching <- shapiro.test(vel_diff_all[Subcomponent == "LtoS"]$Teaching)
vel_diff_ltos_norm_judge <- shapiro.test(vel_diff_all[Subcomponent == "LtoS"]$Mean)

qqnorm(vel_diff_all[Subcomponent == "LtoS"]$Teaching)
qqline(vel_diff_all[Subcomponent == "LtoS"]$Teaching)
vel_diff_ltos_norm_teaching

qqnorm(vel_diff_all[Subcomponent == "LtoS"]$Mean)
qqline(vel_diff_all[Subcomponent == "LtoS"]$Mean)
vel_diff_ltos_norm_judge

cor_vel_diff_ltos <- cor.test(vel_diff_all[Subcomponent == "LtoS"]$Teaching, vel_diff_all[Subcomponent == "LtoS"]$Mean)
cor_vel_diff_ltos
```

## Staccato to Legato (*)
```{r, echo = FALSE}
# normality check
vel_diff_stol_norm_teaching <- shapiro.test(vel_diff_all[Subcomponent == "StoL"]$Teaching)
vel_diff_stol_norm_judge <- shapiro.test(vel_diff_all[Subcomponent == "StoL"]$Mean)

qqnorm(vel_diff_all[Subcomponent == "StoL"]$Teaching)
qqline(vel_diff_all[Subcomponent == "StoL"]$Teaching)
vel_diff_stol_norm_teaching

qqnorm(vel_diff_all[Subcomponent == "StoL"]$Mean)
qqline(vel_diff_all[Subcomponent == "StoL"]$Mean)
vel_diff_stol_norm_judge

cor_vel_diff_stol <- cor.test(vel_diff_all[Subcomponent == "StoL"]$Teaching, vel_diff_all[Subcomponent == "StoL"]$Mean)
cor_vel_diff_stol
```

@@

# KOT
## Legato (n.s.)
### Estimate
- partial correlation coefficient between two variables (Teaching vs KOT)
